\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-27 notes/2loopNLSM.tex
All non-functional comments removed.
Two main comment blocks remain: the tikz-externalize block, and the emacs/auctex options block.

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
%\usepackage{hyperref}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{orcidlink}
\usepackage[normalem]{ulem}

@@

%\usepackage[notcite]{showkeys}
\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={nhpRed},
    citecolor={nhpBlue},
    urlcolor={nhpBlue}
}

@@

%Comment commands
\newcommand{\ace}[1]{\textcolor{darkgreen}{\textbf{AE:}{ #1}}}
\newcommand{\nhp}[1]{\textcolor{nhpRed}{\textbf{NHP:}{#1}}}
\newcommand{\jm}[1]{\textcolor{blue}{\textbf{JM: }{#1}}}

@@

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}
\newcommand{\ansatze}{ans\"atze} %Because words are hard

%TODO-type commands
\newcommand{\citepls}[1]{{\bf\color{red}[[Cite Needed: #1]]}}
\newcommand{\sect}{\textbf{\textcolor{red}{RefSec}}}

@@

%TODO:\\
%For now, let's put the things we need to check/compute here in the abstract. 
%\begin{enumerate}
%\item Settle on $n$ vs $N$ vs $\mathcal{N}$ as the symbol for
 % numerator

%\item \ace{I am starting to get a bit concerned about the overuse of the
  %phrase ``color-dual''} \nhp{We all choose capital $N$}
%\item \ace{I have defined a \textcolor{ucp-color}{\texttt{ucp-color}}
   % , \textcolor{cut-color}{\texttt{cut-color}},
  %  \textcolor{refl-color}{\texttt{refl-color}}, and
  %  \textcolor{jacobi-color}{\texttt{jacobi-color}} as aliases for
  %  other colors so that we don't have to keep changing colors in
   % multiple locations when we want to change diagram color schemes.
    %I think I have updated all of the diagrams to use the
  %  correctly-named color, but haven't changed any of the backing
   % colors for the names.}
%\end{enumerate}
}

@@

%Why DC is cool:
%1) Reduction in computational complexity.
%Off-shell high spin is terrible.
%Gravity has infinitely many very complex Feynman rules as a result of gauge redundancy but YM only has two relatively simple Feynman rules.
%2) DC points to deep structure hidden in Lagrangian formulation.

@@

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows \be % [[11]]
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox + \JacobiOneCrossedBox + \JacobiOnePentaTriangle
=0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic four-point two-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation \be % [[19]]
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} and \cref{BELreg} for discussions of BELs in the context of
color-kinematics duality.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\begin{equation}
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}}\!\! \sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{3}{4}}
\text{ and }
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}} \!\!\sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{4}{3}}
\end{equation}
% \ace{is there a reason we commented out this diagram?} \jm{You can try
 % either one.  I liked the one I left in better.}  For example, the
%symmetries of the crossed box
%\begin{equation}
%\CrossedBoxGraphSym{1}{2}{3}{4}
%\end{equation}
which are just functional relabelings of external legs,
$\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.  In general,
internal edge labels must be tracked as well.  Furthermore, care must
be taken with the signs in a symmetry relation.  To compensate for the
antisymmetric color factors, $f^{abc}$, graph vertices are all totally
antisymmetric and hence have signs built into their orientations.
When used in conjuction with the Jacobi relations, \eg{}
\cref{eq:TwoLoopJacobiOne}, the symmetry constraints of non-basis
graphs still impose conditions on the basis ansatz.  In fact, since
the symmetry properites of a graph are directly linked with
transformations of its color dressing, the resulting constraints can
be thought of as a subclass of the \emph{boundary} Jacobi relations.

@@

%% \jm{Again, I love the colors!
%%  Would it be possible to add $\ell$ labels to the RHS of
%5  \cref{eq:example-cut-sum} to show what $\ell_5$, $\ell_6$... are?}
%%\ace{It is doable, will work on it once our text is a bit more happy}

@@

Even though unitarity methods aim to generate loop-level data from
well-defined on-shell information, the procedure can encounter
subtleties.  Beginning at one loop, there are classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg (BEL)
singularity
\begin{equation}
\frac{\delta(\textcolor{cut-color}{k_4^2})}{s\,\textcolor{ucp-color}{ k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick](i1)--(i2);
    \draw[ultra thick,ucp-color](i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
  \in   \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\end{equation}
in which one of the \textcolor{ucp-color}{uncut propagators} is set
equal to an \textcolor{cut-color}{external on-shell propagator}.
Discussions about how to resolve these types of singularities can be
found in Refs.~\cite{Minahan:1987ha, Bern:2012uf, 
  Geyer:2015bja,Geyer:2015jch,Edison:2020uzf, Edison:2022jln, Edison:2022smn} and \cref{BELreg}. In addition to the
BEL cuts, there is a related class of cuts in which
an uncut propagator is set to zero by an \emph{internal cut
  condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of the
graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
N^{\text{NLSM}}\left[
  \scalebox{0.75}{\NLSMBasisDoubleBox} \right]
\text{ and }
N^{\text{NLSM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]\in
\left\{\begin{array}{cccc}
    k_{13},& k_{23},& k_{15},& k_{16}, \\
    k_{25},& k_{26},& k_{35},& k_{36}, \\
    k_{55},& k_{56},& k_{66}
  \end{array}
  \right\},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
%%%%
% Basis for both diagrams
% {d[p[1], p[3]], d[p[1], p[6]], d[p[1], p[7]], d[p[2], p[3]],  d[p[2], p[6]], d[p[2], p[7]], d[p[3], p[6]], d[p[3], p[7]],  d[p[6], p[6]], d[p[6], p[7]], d[p[7], p[7]]}
%%%%
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to solving large sets of equations with sixteen thousand
  parameters, so a different method is desirable.  One possibility is
  to use a sprase solver, as often employed for integral reduction.
  However, these solvers typically assume that the matrix has a very
  low density.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions) which limits the
  cross-talk between equations and thus the effective density, there
  is no such separation of sectors in a scalar theory.  Thus the NLSM
  constraint equations were solved using a custom solver, working name
  \texttt{FiniteFieldSolve}, designed to exactly solve large linear
  systems of arbitrary density.  \texttt{FiniteFieldSolve} will be
  released publicly shortly.} Imposing the boundary Jacobi relations
eliminates all but 4,473 of the original free parameters.  Further
imposing graph symmetries reduces the number of free parameters to
1,243.

@@

% It is possible to automatically enforce (or ``bake in'') the double
% box and penta-triangle max cuts on the ansatz by writing the ansatz
% in terms of inverse propagators (and remaining irreducible scalar
% products) and then only keeping terms that involve at least one
% inverse propagator since those terms will vanish on those max cuts,
% just as the product of 3pt scalar amplitudes will.  However, for the
% sake of simplicity we do not take this route.

@@

\begin{figure}[t]
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.
The numerators of the two basis graphs are
%\left[\begin{split}
%  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_2+k_1+k_2) \rangle 
%\\
%\langle (l_1+k_3+k_4)(l_2+k_1+k_2) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
%\end{split}\right]
%
%\left[\begin{split}
%  \langle (l_1+k_4) l_1 \rangle\langle l_1l_2 \rangle\langle l_2 (l_2+k_1) \rangle  \langle (l_2+k_1)(l_1-k_1) \rangle 
%\\
%\langle (l_1+k_3+k_4)(l_1-k_1) \rangle\langle (l_1+k_3+k_4)(l_1+k_4) \rangle
%\end{split}\right]
%
\begin{align}
N^{\text{ZM}}\left[ \scalebox{0.75}{\ZMBasisDoubleBox} \right] &=
\begin{aligned}
&\langle \ell_1 \ell_2 \rangle \langle k_4 \ell_1 \rangle \langle \ell_2 k_1 \rangle \langle k_3 (\ell_1+k_4)\rangle \\
& \times \langle (k_1+\ell_2) k_2 \rangle \langle (\ell_2+k_{12})(\ell_1-k_{12})\rangle
\end{aligned} \\
N^{\text{ZM}}\left[ \scalebox{0.75}{\ZMBasisPentaTriangle} \right] &=
\begin{aligned}
&\langle \ell_1 \ell_2 \rangle \langle k_4 \ell_1 \rangle \langle \ell_2 k_1 \rangle \langle k_3 (\ell_1+k_4) \rangle \\
& \times  \langle (k_1-\ell_1) k_2 \rangle \langle (\ell_2+k_1) (\ell_1-k_1)\rangle
\end{aligned}
\end{align}
where $k_{12} = k_1+k_2$ and, again, $\langle ab\rangle \equiv p_a^\mu \varepsilon_{\mu\nu} p_b^\nu$ is the color-stripped ZM vertex.
As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.
We introduce a proportionality constant $z$ between the numerators of the two theories, $N_\text{NLSM} \vert_{2D} = z ~ N_\text{ZM}$, because we are only interested in how the kinematical structure of $N_\text{ZM}$ can be used to eliminate gauge freedom in $N_\text{NSLM}$.\footnote{Another reason for introducing the parameter $z$ is because the ZM and NLSM are believed to differ at loop level even though they are dual classically in 2D \cite{Nappi:1979ig}.  Once all of the constraints from this section are imposed, $z$ is fixed to $216/565$ where the four-point pion tree amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$.}
Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.

>>>>

\documentclass[11pt,letter]{article}
\pdfoutput=1
\usepackage{jheppub}
\usepackage{amsmath,amssymb}
\usepackage[dvipsnames]{xcolor}
\usepackage{xspace}
\usepackage{ifdraft}
\usepackage{epstopdf}
\usepackage{slashed}
\usepackage{anyfontsize}
\usepackage{tocloft}
\renewcommand{\cftdot}{.}
\renewcommand{\cftsubsecleader}{\cftdotfill{\cftdotsep}}
\usepackage{diagbox}
\usepackage{colortbl}
\usepackage{bbold}
\usepackage{tabularx}
\usepackage{graphicx}
\usepackage{stackengine}
\usepackage{subcaption}
\usepackage{cleveref}
\usepackage{orcidlink}
\usepackage[normalem]{ulem}

@@

\usepackage{tikz-feynman} 
\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={nhpRed},
    citecolor={nhpBlue},
    urlcolor={nhpBlue}
}

@@

@@

\def\tree{{\rm tree}}
\def\pol{\varepsilon}
\def\Tr{\, {\rm Tr}}
\def\tr{\, {\rm tr}}
\def\eps{\varepsilon}
\def\e{\varepsilon}
\def\ep{\varepsilon}
\def\SYM{MSYM}
\def\nn{\nonumber}
\def\Neqfour{{{\cal N}=4}}
\def\NeqFour{{{\cal N}=4}}
\def\Neqeight{{{\cal N}=8}}
\def\NeqEight{{{\cal N}=8}}
\def\NeqOne{{{\cal N}=1}}
\def\Fact{{\cal F}}
\def\f{\widetilde f}
\def\be{\begin{equation}}
\def\ee{\end{equation}}
\def\bea{\begin{eqnarray}}
\def\eea{\end{eqnarray}}
\def\ba{\begin{eqnarray}}
\def\ea{\end{eqnarray}}
\def\Ksl{{\s K}}
\def\ksl{\s{k}}
\def\Perm{{\cal P}}
\def\M{{\cal M}}
\def\ve{\varepsilon}
\def\tlambda{{\tilde\lambda}}
\def\MHVbar{$\overline{\hbox{MHV}}$}
\def\NMHVbar{$\overline{\hbox{NMHV}}$}
\def\P{{\rm P}}
\def\NHP{{\rm NHP}}
\def\mud{\lambda}
\def\bowtie{{\rm bow\mbox{-}tie}}
\newcommand{\cred}{\bf \color{red}}
 \newcommand{\cblue}{\color{blue}}
\newcommand{\andd}{\ , \quad \text{and}  \quad}
\newcommand{\forr}{\ , \quad \text{for}  \quad}
\newcommand{\ansatze}{ans\"atze}

@@

}

@@

@@

With the integrand expressed in terms of cubic graphs, it is possible
to discuss the main ingredient in color-kinematics duality -- the
Jacobi relations.  The color factors in an integrand obey a set of
Jacobi identities.  Color-kinematics duality states that there
\emph{exists} a way of writing the kinematic numerators of the
integrand so that they obey the same set of Jacobi identities.  The
Jacobi identities can be divided into two categories: \emph{defining}
relations and \emph{boundary} relations.  Defining Jacobi relations
can be used to express any graph in terms of a basis.  For example,
the double box, crossed box, and penta-triangle are related to each
other as follows
\begin{equation}
\label{eq:TwoLoopJacobiOne}
\JacobiOneDoubleBox + \JacobiOneCrossedBox + \JacobiOnePentaTriangle
=0\,,
\end{equation}
where the shared propagator is shown in red.  Using this relation the
crossed box can be written (or \emph{defined}) in terms of the other
two.  In this way, all of the 14 cubic four-point two-loop topologies can be
written in terms of a basis of the double box and penta-triangle.
(Any two of the double box, crossed box, and penta-triangle would
suffice for a basis but the planar basis is chosen for simplicity in
this work.)  The numerator of each of the two basis graphs receives
its own ansatz and the numerator of any other graph can be expressed
in terms of these two.

After using the defining Jacobi relations to relate every graph to the
basis, any remaining Jacobi relations will be referred to as boundary
relations.  In other words, boundary relations (indirectly) relate
basis elements to themselves.  For example, the crossed box numerator
is related to itself via the following Jacobi relation
\begin{equation}
\label{eq:TwoLoopJacobiTwo}
\JacobiTwoCrossedBoxOne + \JacobiTwoCrossedBoxTwo +  \JacobiTwoCrossedBoxThree =0,
\end{equation}
where the shared propagator is shown in red as before.  Each of the
crossed box numerators can be written in terms of the double box and
penta-triangle so this Jacobi relation places constraints on the basis
numerator \ansatze{}.  While defining Jacobi relations simplify how
many numerators require \ansatze{}, boundary relations place
restrictions on those \ansatze{}.  In principle there could be Jacobi
relations involving tadpoles and BEL graphs, which would be subtle to
handle.  One option is to force the numerators of these graphs to
zero.  However, since these graphs will not contribute to physical
unitarity cuts, we take a different approach.  These graphs are
allowed to have non-zero numerators but any Jacobi relations involving
them are simply not imposed or solved.  See Refs.~\cite{Bern:2012uf,
  Edison:2022jln} and \cref{BELreg} for discussions of BELs in the context of
color-kinematics duality.

@@

Every numerator is required to respect all of the symmetries (or
automorphisms) of its associated graph.  Graph symmetries are an
avatar of Bose symmetry.  For a pure ansatz construction it is very
reasonable to impose graph symmetries, but color-kinematics duality at
tree-level is often achieved at the expense of manifest Bose symmetry
\cite{Cheung:2016prv, Cheung:2021zvb, Brandhuber:2021bsf,
  Edison:2020ehu}.  In an ansatz construction graph symmetries are
imposed for \emph{every} graph, not just the basis graphs.
For example, the symmetries of the crossed box are spanned by,
\begin{equation}
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}}\!\! \sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{3}{4}}
\text{ and }
\scalebox{0.8}{\CrossedBoxGraphSym{1}{2}{3}{4}} \!\!\sim\!\! \scalebox{0.8}{\CrossedBoxGraphSym{2}{1}{4}{3}}
\end{equation}
which are just functional relabelings of external legs,
$\{1234\} \to \{2134\}$ and $\{1234\} \to \{1243\}$.  In general,
internal edge labels must be tracked as well.  Furthermore, care must
be taken with the signs in a symmetry relation.  To compensate for the
antisymmetric color factors, $f^{abc}$, graph vertices are all totally
antisymmetric and hence have signs built into their orientations.
When used in conjuction with the Jacobi relations, \eg{}
\cref{eq:TwoLoopJacobiOne}, the symmetry constraints of non-basis
graphs still impose conditions on the basis ansatz.  In fact, since
the symmetry properites of a graph are directly linked with
transformations of its color dressing, the resulting constraints can
be thought of as a subclass of the \emph{boundary} Jacobi relations.

@@

@@

Even though unitarity methods aim to generate loop-level data from
well-defined on-shell information, the procedure can encounter
subtleties.  Beginning at one loop, there are classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg (BEL)
singularity
\begin{equation}
\frac{\delta(\textcolor{cut-color}{k_4^2})}{s\,\textcolor{ucp-color}{ k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick](i1)--(i2);
    \draw[ultra thick,ucp-color](i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
  \in   \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\end{equation}
in which one of the \textcolor{ucp-color}{uncut propagators} is set
equal to an \textcolor{cut-color}{external on-shell propagator}.
Discussions about how to resolve these types of singularities can be
found in Refs.~\cite{Minahan:1987ha, Bern:2012uf, 
  Geyer:2015bja,Geyer:2015jch,Edison:2020uzf, Edison:2022jln, Edison:2022smn} and \cref{BELreg}. In addition to the
BEL cuts, there is a related class of cuts in which
an uncut propagator is set to zero by an \emph{internal cut
  condition}, for instance,
\begin{equation}
  \label{eq:other-path}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

@@

For a scalar theory like NLSM the numerator only depends on dot
products of momenta.  The basis of momentum invariants for both of the
graphs can be taken to be
\begin{equation}
\label{eq:scalar-basis}
N^{\text{NLSM}}\left[
  \scalebox{0.75}{\NLSMBasisDoubleBox} \right]
\text{ and }
N^{\text{NLSM}}\left[\scalebox{0.75}{\NLSMBasisPentaTriangle} \right]\in
\left\{\begin{array}{cccc}
    k_{13},& k_{23},& k_{15},& k_{16}, \\
    k_{25},& k_{26},& k_{35},& k_{36}, \\
    k_{55},& k_{56},& k_{66}
  \end{array}
  \right\},
\end{equation}
where $k_{ij}$ means $k_i \cdot k_j$, $k_5=\ell_1$, and $k_6 = \ell_2$.
Every vertex in NLSM scales as $k^2$ so the numerator of each diagram
scales as $k^{12}$.  At this mass dimension there are 8,008
independent dot products of momenta.  With two basis graphs there is a
total of 16,016 ansatz parameters.\footnote{\texttt{Mathematica} is
  not suited to solving large sets of equations with sixteen thousand
  parameters, so a different method is desirable.  One possibility is
  to use a sprase solver, as often employed for integral reduction.
  However, these solvers typically assume that the matrix has a very
  low density.  While the equations in YM naturally separate into
  different ``helicity'' sectors (or $e_i \cdot e_j$ and
  $e_i \cdot k_j$ sectors in $D$ dimensions) which limits the
  cross-talk between equations and thus the effective density, there
  is no such separation of sectors in a scalar theory.  Thus the NLSM
  constraint equations were solved using a custom solver, working name
  \texttt{FiniteFieldSolve}, designed to exactly solve large linear
  systems of arbitrary density.  \texttt{FiniteFieldSolve} will be
  released publicly shortly.} Imposing the boundary Jacobi relations
eliminates all but 4,473 of the original free parameters.  Further
imposing graph symmetries reduces the number of free parameters to
1,243.

@@

@@

\begin{figure}[t]
  \begin{center}
	\PhysicalCutOne{}{}{}{} \PhysicalCutTwo{}{}{}{}
  \end{center}
  \caption{The two physical cut topologies shared by NLSM, sGal, BI, and DBIVA}
  \label{fig:emu}
\end{figure}
It is tempting to try to match the pion four-point two-loop integrand to ZM
on the cuts, but given that the cuts only probe the on-shell four-point
amplitude, which is either subtle or vanishes for ZM, it is better to
compare the off-shell numerators directly.
The numerators of the two basis graphs are
\begin{align}
N^{\text{ZM}}\left[ \scalebox{0.75}{\ZMBasisDoubleBox} \right] &=
\begin{aligned}
&\langle \ell_1 \ell_2 \rangle \langle k_4 \ell_1 \rangle \langle \ell_2 k_1 \rangle \langle k_3 (\ell_1+k_4)\rangle \\
& \times \langle (k_1+\ell_2) k_2 \rangle \langle (\ell_2+k_{12})(\ell_1-k_{12})\rangle
\end{aligned} \\
N^{\text{ZM}}\left[ \scalebox{0.75}{\ZMBasisPentaTriangle} \right] &=
\begin{aligned}
&\langle \ell_1 \ell_2 \rangle \langle k_4 \ell_1 \rangle \langle \ell_2 k_1 \rangle \langle k_3 (\ell_1+k_4) \rangle \\
& \times  \langle (k_1-\ell_1) k_2 \rangle \langle (\ell_2+k_1) (\ell_1-k_1)\rangle
\end{aligned}
\end{align}
where $k_{12} = k_1+k_2$ and, again, $\langle ab\rangle \equiv p_a^\mu \varepsilon_{\mu\nu} p_b^\nu$ is the color-stripped ZM vertex.
As local functions, the
numerators never suffer from any of the subtleties of on-shell 2D
kinematics.
We introduce a proportionality constant $z$ between the numerators of the two theories, $N_\text{NLSM} \vert_{2D} = z ~ N_\text{ZM}$, because we are only interested in how the kinematical structure of $N_\text{ZM}$ can be used to eliminate gauge freedom in $N_\text{NSLM}$.\footnote{Another reason for introducing the parameter $z$ is because the ZM and NLSM are believed to differ at loop level even though they are dual classically in 2D \cite{Nappi:1979ig}.  Once all of the constraints from this section are imposed, $z$ is fixed to $216/565$ where the four-point pion tree amplitude is normalized to $-k_1\cdot k_2$ and the ZM vertex is normalized to $k_1^\mu \varepsilon_{\mu\nu}k_2^\nu$.}
Mechanically, the pion numerator is matched
to ZM by first taking every possible assignment of right and left
movers for the external particles and then restricting the pion
numerator to 2D.  The loop momenta are restricted to 2D but left
off-shell.  After performing the 2D matching, there are only 365
parameters of generalized gauge freedom.
